### 2019-03-16 paper/listing.tex

\usepackage{graphicx}
\usepackage{xspace}
%\usepackage{url}
\usepackage{listings}
% The following packages are used in the Scheme definition

\lstdefinelanguage{java}{
morekeywords={%Java keywords
float, public, interface, class, static, void, extends, implements, final,
boolean, return, new, abstract, super, for, int, package, private, import,
protected, this, throw, try, finally, if, else, while, instance of, &&, ||},
	sensitive=true,
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/},
	morestring=[b]",
}

\lstdefinelanguage{tracematches}
{morekeywords={%Java keywords
float, public, interface, class, static, void, extends, implements, final,
boolean, return, new, abstract, super, for, int, package, private, import,
protected, this, throw, try, finally, if, else, while, instance of, &&, ||,
%Aspectj keywords
before, around, after, returning, tracematch, sym, call, exec},
	sensitive=true,
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/},
	morestring=[b]",
}

\lstdefinelanguage{JavaScript}
	{morekeywords={%Java keywords
	               true, false, return, new, function, var,
	               this, throw, try, finally, if, else, while, for, instanceof, 
	               %Others
	               async, await},
	sensitive=true,
	morecomment=[l]{//},
	morecomment=[s]{/*}{*/},
	morestring=[b]",
}

\lstset{
	frame=,
	language=JavaScript,
	xleftmargin=6pt,
	stepnumber=1,
	numbers=none,
	numbersep=5pt,
	numberstyle=\ttfamily\tiny,
	belowcaptionskip=\bigskipamount,
	captionpos=b,
	escapeinside={*'}{'*},
	tabsize=2,
	emphstyle={\bf},
	stringstyle=\mdseries\rmfamily,
	showspaces=false,
	keywordstyle=\bfseries,
    basewidth=0.5em,
	columns=fixed,
	%basicstyle=\footnotesize\sffamily,
         basicstyle=\scriptsize,
	showstringspaces=false,
	morecomment=[l]\%,
	commentstyle=\it
}

%\newcommand{\co}[1]{\mbox{{\small #1}\xspace}}
\newcommand{\co}[1]{\mbox{{\fontfamily{phv}\selectfont\scriptsize #1}\xspace}}
\newcommand{\env}[1]{\framebox{{\fontfamily{phv}\selectfont\scriptsize
#1}\xspace}}
\newcommand{\lambdac}[0]{$\lambda$~}
%\renewcommand{\ttdefault}{cmtt}
%\renewcommand{\ttdefault}{txtt}